---
title: "ECON 573 Problem Set 1 Part 2 and 3"
output:
  pdf_document: default
  html_notebook: default
---

## Part 2

8a) Use the read.csv() function to read the data into R. Call the
loaded data college. Make sure that you have the directory set
to the correct location for the data.
```{r}
college = read.csv("College.csv")
```

8b) Look at the data using the fix() function. You should notice
that the first column is just the name of each university.We don’t
really want R to treat this as data. However, it may be handy to
have these names for later. Try the following commands:
```{r}
rownames(college) = college[,1]
college = fix(college)
college = college [,-1]
college = fix(college)

```

8ci) Use the summary() function to produce a numerical summary
of the variables in the data set.
```{r}
summary(college)
```

8cii). Use the pairs() function to produce a scatterplot matrix of
the first ten columns or variables of the data. Recall that
you can reference the first ten columns of a matrix A using
A[,1:10].
```{r}
college[,1] = as.numeric(factor(college[,1]))
pairs(college[, 1:10])
```

8ciii) Use the plot() function to produce side-by-side boxplots of
Outstate versus Private.
```{r}
attach(college)
boxplot(Outstate ~ Private, xlab = "Private", ylab = "Outstate")
```

8civ) Create a new qualitative variable, called Elite, by binning
the Top10perc variable. We are going to divide universities
into two groups based on whether or not the proportion
of students coming from the top 10% of their high school
classes exceeds 50%.
```{r}
Elite = rep("No",nrow(college ))
Elite [college$Top10perc >50]=" Yes"
Elite = as.factor (Elite)
college = data.frame(college ,Elite)
```
Use the summary() function to see how many elite universities
there are. Now use the plot() function toproduce
side-by-side boxplots of Outstate versus Elite.
```{r}
summary(college$Elite)
boxplot(college$Outstate ~ college$Elite, xlab = "Elite", ylab = "Outstate")
```

8cv) Use the hist() function to produce some histograms with
differing numbers of bins for a few of the quantitative variables.
You may find the command par(mfrow=c(2,2)) useful:
it will divide the print window into four regions so that four
plots can be made simultaneously. Modifying the arguments
to this function will divide the screen in other ways.
```{r}
par(mfrow=c(2,2))
hist(college$S.F.Ratio, breaks = 5)
hist(college$perc.alumni, breaks = 6)
hist(college$Grad.Rate, breaks = 8)
hist(college$P.Undergrad, breaks = 4)
```

8cvi) Continue exploring the data, and provide a brief summary
of what you discover.

9) This exercise involves the Auto data set studied in the lab. Make sure
that the missing values have been removed from the data.
```{r}
auto = read.csv("Auto.csv")
auto = na.omit(auto)
detach(college)
attach(auto)
```

9a) Which of the predictors are quantitative, and which are qualitative?
```{r}
sapply(auto, class)
```
From the output of the command it seems as if name is the only qualitative variable

9b) What is the range of each quantitative predictor? You can answer
this using the range() function.
```{r}
apply(auto[,1:7], 2, range)
```

9c) What is the mean and standard deviation of each quantitative
predictor?
```{r}
sapply(auto[,1:7], mean, na.rm = TRUE)
sapply(auto[,1:7], sd, na.rm = TRUE)
```

9d) Now remove the 10th through 85th observations. What is the
range, mean, and standard deviation of each predictor in the
subset of the data that remains?
```{r}
sapply(auto[-c(10:85),1:7], range, na.rm = TRUE)
sapply(auto[-c(10:85),1:7], mean, na.rm = TRUE)
sapply(auto[-c(10:85),1:7], sd, na.rm = TRUE)
```

9e) Using the full data set, investigate the predictors graphically,
using scatterplots or other tools of your choice. Create some plots
highlighting the relationships among the predictors. Comment
on your findings.
```{r}
auto$horsepower <- as.numeric(type.convert(auto$horsepower))
pairs(auto[,1:7])
```

9f) Suppose that we wish to predict gas mileage (mpg) on the basis
of the other variables. Do your plots suggest that any of the
other variables might be useful in predicting mpg? Justify your
answer.
```{r}
regression = lm(mpg ~ cylinders+year+weight+horsepower+displacement+acceleration+origin, data= auto)
regression
```

10) This exercise involves the Boston housing data set.
```{r}
library(MASS)
data(Boston)
```

10a) How many rows are in this data set? How many columns? What
do the rows and columns represent?
506 rows, 14 columns
```{r}
dim(Boston)
```

10b) Make some pairwise scatterplots of the predictors (columns) in
this data set. Describe your findings.
looking att the pairs theres a large jumble of data, some of the variable are definitely coorelated
while others seem to show nothing at all
```{r}
pairs(Boston)
```

10c) Are any of the predictors associated with per capita crime rate?
If so, explain the relationship.
It seems that variables with the strongest correlation coefficient are possitive
```{r}
Boston.corr.crim = cor(Boston)[-1,1]
print(Boston.corr.crim[order(abs(Boston.corr.crim))])
```

10d) Do any of the suburbs of Boston appear to have particularly
high crime rates? Tax rates? Pupil-teacher ratios? Comment on
the range of each predictor.
-About 15 sububurs have high crime rates
-There seems to be one outlier in the data, but otherwise tax is uniform
-There seems to be higher skew in teacher ratios
```{r}
par(mfrow=c(1,3))
hist(Boston$crim[Boston$crim > 1], breaks=15)
hist(Boston$tax, breaks=15)
hist(Boston$ptratio, breaks=15)
```

10e) How many of the suburbs in this data set bound the Charles
river?
```{r}
sum(Boston$chas == 1)
```

10f) What is the median pupil-teacher ratio among the towns in this
data set?
```{r}
median(Boston$ptratio)
```

10g) Which suburb of Boston has lowest median value of owneroccupied
homes? What are the values of the other predictors
for that suburb, and how do those values compare to the overall
ranges for those predictors? Comment on your findings.
```{r}
t(subset(Boston, medv == min(medv)))
```

10h) In this data set, how many of the suburbs average more than
seven rooms per dwelling? More than eight rooms per dwelling?
Comment on the suburbs that average more than eight rooms
per dwelling.
```{r}
sum(Boston$rm > 7)
sum(Boston$rm > 8)
summary(subset(Boston, rm > 8))
```

## Part 3

These lines of code are from the starter file!
```{r}
benjer = read.csv("BenAndJerry.csv")
names(benjer)
priceper1 = (benjer$price_paid_deal + benjer$price_paid_non_deal)/benjer$quantity
y <- log(1+priceper1)
x <- benjer[,c("flavor_descr","size1_descr",
	"household_income","household_size")]
x$flavor_descr <- relevel(factor(x$flavor_descr),"VAN")
x$usecoup = factor(benjer$coupon_value>0)
x$couponper1 <- benjer$coupon_value/benjer$quantity
x$region <- factor(benjer$region, 
	levels=1:4, labels=c("East","Central","South","West"))
x$married <- factor(benjer$marital_status==1)
x$race <- factor(benjer$race,
	levels=1:4,labels=c("white","black","asian","other"))
x$hispanic_origin <- benjer$hispanic_origin==1
x$microwave <- benjer$kitchen_appliances %in% c(1,4,5,7)
x$dishwasher <- benjer$kitchen_appliances %in% c(2,4,6,7)
x$sfh <- benjer$type_of_residence==1
x$internet <- benjer$household_internet_connection==1
x$tvcable <- benjer$tv_items>1
fit <- glm(y~., data=x) 
pvals <- summary(fit)$coef[-1,4]  
par(mfrow=c(1,1))
par(mar=c(5,10,5,5))
barplot(table(benjer$flavor),border=NA, las=2)
class(benjer$promotion_type)
barplot(table(benjer$flavor),border=NA, las=2)
hist(y); hist(priceper1);
hist(benjer$promotion_type)
hist(benjer$household_income)
hist(benjer$age_of_female_head)
hist(benjer$age_of_male_head)
hist(benjer$race)
num_cols <- unlist(lapply(benjer, is.numeric))
boxplot(y[benjer$promotion_type>0] ~ benjer$promotion_type, col= levels(factor(benjer$promotion)))
summary(fit)
fit2 <- glm(y~. - flavor_descr, data=x) 
summary(fit2)
fit3 <- glm(y~. -flavor_descr -race -dishwasher - hispanic_origin, data=x)
summary(fit3)
fit4 <- glm(y~. -flavor_descr -race -dishwasher - hispanic_origin -internet -tvcable, data=x)
summary(fit4)
n=nrow(x)
BIC <- c(reg1=extractAIC(fit, k=log(n))[2],
         reg2=extractAIC(fit2, k=log(n))[2],
         reg3=extractAIC(fit3, k=log(n))[2],
         reg4=extractAIC(fit4, k=log(n))[2])
eBIC <- exp(-0.5*(BIC-min(BIC)))
probs <- eBIC/sum(eBIC)
round(probs, 5)
```
1. Explore the data and visualize: what variables are interesting? Choose a few, plot
them together, and tell a story.
As can be seen from the dataset the age of heads seems to have a large significan amount at the younger
population as it steadliy increases throughout the age group.Those who do eat at ben and jerry's also
tend to have an higher household income. Furthermore the majority of people tend to of race type 1.
There also seems to be a higher spread among promotions of type 3

2. Describe the regression model in the code. Improve it?
The first model is a baseline in which the log of price per ice cream is regressed against many variables.
When the flavor variable was taken out it was found that, asian, hispanic, and dishwashwer, were not
significant so they were taken out of the model, which is the third one. Perhaps one way to improve the
model is to only include regressors that are significant at the 0.01 alpha level. If this is done, we remove
the varialbe internet and tvcable. Doing so will show a higher BIC value for this model. This can be seen
in the code by fit4

3. Take the p-values from your regression and look for evidence of association. Relate
what you learn to your story from 1.
From the p-values it can be seen that all aggressors in fit4 are statistically signifcant at the 0.1
alpha level. Thus the variables of size1_descr32.0 MLOZ  0.3121873  0.0077971  40.039  < 2e-16 ***
household_income, household_size, usecoupTRUE, couponper1, regionCentral, regionSouth,regionWest, 
marriedTRUE, microwaveTRUE, and sfhTRUE do seem to have an association on the log of price.
